### 2024-06-04 draft.qmd
20240603 - fix headings for cross-references
20240603 - clarification
20240604 - more player info

### Auction Draft [#sec-draftAuction]

@@

An important concept in the draft is "dropoff."
*Dropoff* at a given position, is the difference—in terms of projected fantasy points—between (a) the best available player remaining at that position and (b) the second-best available player remaining at that position.
If there is a bigger dropoff at a given position, there may be greater value in drafting the top player from that position.
For instance, consider the following scenario:
"Quarterback A" is projected to score 325 points, and "Quarterback B" is projected to score 320 points.
"Tight End A" is projected to score 230 points, and "Tight End B" is projected to score 150 points.
In this example, there is a much greater dropoff for Tight Ends than there is for Quarterbacks.
Thus, even though "Quarterback A" is projected to score more points than "Tight End A", "Tight End A" may be more valuable because there is still a good Quarterback available if someone else drafts "Quarterback A".

@@

An important concept in the snake draft is a player's value over a typical replacement player (shortened to "value over replacement player").
A player's *value over a replacement player* (VORP) is the difference between (a) a player's projected fantasy points and (b) the fantasy points that you would be expected to get from a typical bench player at that position.

### Auction Draft [#sec-draftStrategyAuction]

>>>>

### Auction Draft {#sec-draftAuction}

@@

In general, you will first want to generate the rankings you will use to select which players to prioritize.
You may generate your rankings based one or more of the following:

    - your [player evaluation](player-evaluation.qmd)
    - expert or aggregated rankings
    - players' Average Draft Position (ADP) in other league drafts (for snake drafts), or players' Average Auction Value (AAV) in other league drafts (for auction drafts)
    - expert or aggregated projections

An important concept in the draft is "dropoff", which is described in Section @sec-fantasyValue.
*Dropoff* at a given position, is the difference—in terms of projected fantasy points—between (a) the best available player remaining at that position and (b) the second-best available player remaining at that position.
If there is a bigger dropoff at a given position, there may be greater value in drafting the top player from that position.
For instance, consider the following scenario:
"Quarterback A" is projected to score 325 points, and "Quarterback B" is projected to score 320 points.
"Tight End A" is projected to score 230 points, and "Tight End B" is projected to score 150 points.
In this example, there is a much greater dropoff for Tight Ends than there is for Quarterbacks.
Thus, even though "Quarterback A" is projected to score more points than "Tight End A", "Tight End A" may be more valuable because there is still a good Quarterback available if someone else drafts "Quarterback A".

@@

An important concept in the snake draft is a player's value over a typical replacement player at that position (shortened to "value over replacement player"), which is described in Section @sec-fantasyValue.
A player's *value over a replacement player* (VORP) is the difference between (a) a player's projected fantasy points and (b) the fantasy points that you would be expected to get from a typical bench player at that position.

### Auction Draft {#sec-draftStrategyAuction}

### 2024-06-04 player-evaluation.qmd
20240604 - more player info

- athletic profile
- historical performance
- health
- age and career stage
- situational factors
- matchups
- cognitive and motivational factors

@@

In addition to how many fantasy points a player scores in terms of historical performance, we also care about efficiency.
How efficient were they given the number of opportunities they had?
If they were relatively more efficient, they will likely score more points than many of their peers when given more opportunities.
If they were relativelly inefficient, their capacity to score fantasy points may be more dependent on touches/opportunities.
Efficiency might be operationalized by indicators such as yards per passing attempt, yards per rushing attempt, yards per target, yards per reception, etc.

@@

@@

After performing an evaluation of the relevant domain(s) for a given player, then one must integrate the evaluation information across domains to make a judgment about a player's overall value.
When thinking about a player's value, it can be worth thinking of a player's upside (or ceiling) and a player's downside (or flow).
Player that are more consistent may show higher downside but a lower upside.
Younger, less experienced players may show a higher upside but a lower downside.

>>>>

- athletic profile
- historical performance
- health
- age and career stage
- situational factors
- matchups
- cognitive and motivational factors
- fantasy value

@@

In addition to how many fantasy points a player scores in terms of historical performance, we also care about efficiency and consistency.
How efficient were they given the number of opportunities they had?
If they were relatively more efficient, they will likely score more points than many of their peers when given more opportunities.
If they were relativelly inefficient, their capacity to score fantasy points may be more dependent on touches/opportunities.
Efficiency might be operationalized by indicators such as yards per passing attempt, yards per rushing attempt, yards per target, yards per reception, etc.

In terms of consistency, how consistent was the player they from game to game and from season to season?
For instance, we could examine the standard deviations of players' fantasy points across games in a given season.
However, the standard deviation tends to be upwardly biased as the mean increases.
So, we can account for the player's mean fantasy points per game by dividing their game-to-game standard deviation of fantasy points by their mean fantasy points across games.
This is known as the coefficient of variation (CV):

$$
CV = \frac{\sigma}{\mu}
$$

Players with a lower standard deviation and a lower coefficient of variation (of fantasy points across games) are more consistent.

[ADD EXAMPLE]

@@

## Fantasy Value {#sec-fantasyValue}

There are several sources that one can draw upon to evaluate a player's fantasy value:

- expert or aggregated rankings
- players' Average Draft Position (ADP) in other league drafts
- players' Average Auction Value (AAV) in other league drafts
- expert or aggregated projections

[ADD LINKS TO SOURCES OF RANKINGS/PROJECTIONS/ADP/AAV, ETC]

It is important to keep in mind that rankings, average draft position (ADP), and average auction values (AAV) are specific to roster and scoring settings of a particular league.
For instance, [DESCRIBE]
Thus, if using rankings, ADP, or AAV, it is important to find ones from leagues that mirror—as closely as possible—your league settings.

Projected statistics (e.g., projected passing touchdowns) are agnostic to league settings and can thus be used to generate league-specific fantasy projections and rankings.
Thus, projected statisitics may be more useful than rankings because they can be used to generate rankings for your particular league settings.
For instance, if you know how many touchdowns, yards, and interceptions a Quarterback is a projected to throw (in addition to any other relevant categories for the player, e.g., rushing yards and touchdowns), you can calculate how many fantasy points the Quarterback is expected to gain in your league (or in any league).
Thus, you can calculate ranking from projections, but you cannot reverse engineer projections from rankings.

There are many potential indices that can be worthwhile to consider, including a player's:

- dropoff
- value over replacement player (VORP)
- uncertainty

A player's *dropoff*, is the difference between (a) the player's projected points and (b) the projected points of the next-best player at that position.
We can also consider a player's value over a typical replacement player at that position (shortened to "value over replacement player").
A player's *value over a replacement player* (VORP) is the difference between (a) a player's projected fantasy points and (b) the fantasy points that you would be expected to get from a typical bench player at that position.

A player's *uncertainty* is how much variability there is in projections or rankings for a given player across sources.
For instance, consider a scenario where three experts provide ratings about two players, Player A and Player B.
Player A is projected to score 300, 310, and 290 points by experts 1, 2, and 3, respectively.
Player B is projected to score 400, 300, and 200 points by experts 1, 2, and 3, respectively.
In this case, both players are (on average) projected to score the same number of points (300).

```{r}
exampleData <- data.frame(
  player = c(rep("A", 3), rep("B", 3)),
  expert = c(1:3, 1:3),
  projectedPoints = c(300, 310, 290, 400, 300, 200)
)

playerA_mean <- mean(exampleData$projectedPoints[which(exampleData$player == "A")])
playerB_mean <- mean(exampleData$projectedPoints[which(exampleData$player == "B")])

playerA_mean
playerB_mean
```

However, the players differ considerably in their uncertainty (i.e., the source-to-source variability in their projections), as operationalized with the standard deviation and coefficient variation of projected points across sources for a given player.

```{r}
playerA_sd <- sd(exampleData$projectedPoints[which(exampleData$player == "A")])
playerb_sd <- sd(exampleData$projectedPoints[which(exampleData$player == "B")])

playerA_cv <- playerA_mean / playerA_sd
playerB_cv <- playerB_mean / playerb_sd

playerA_mean
playerB_mean

playerA_sd
playerb_sd

playerA_cv
playerB_cv
```

Uncertainty is not necessarily a bad.
It just means we have less confidence about how they may be expected to perform.
Thus, players with greater uncertainty are risky and tend to have a higher upside (or ceiling) and a lower downside (or floor).

@@

After performing an evaluation of the relevant domain(s) for a given player, then one must integrate the evaluation information across domains to make a judgment about a player's overall value.
When thinking about a player's value, it can be worth thinking of a player's upside and a player's downside.
Player that are more consistent may show higher downside but a lower upside.
Younger, less experienced players may show a higher upside but a lower downside.
